---
title: "ESM 206 Lab 5"
author: "David Segan"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridate} package to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}

library(tidyverse)
library(here)
library(janitor)

```


```{r}

# Import lobster data

lobster_abundance <- read_csv(here("data","lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()

```

Use the tidyr::uncount() function to expand a frequency table. Expands observation for as many counts there are listed. Gets rid of count column.

```{r}

lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

Yayayay! Now each lobster has its own row. This is tidy format.

Now let's look at this a bunch of ways.

```{r}

ggplot(data = lobster_tidy,
       aes(x = site,
           y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)

ggplot(data = lobster_tidy,
       aes(x = site,
           y = size_mm)) +
  geom_violin(aes(color = site),
              width = 0.6,
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill=site)) +
  facet_wrap(~site, scales = "free") # scales = free means individual scales can differ based off maxes and mins

ggplot(data = lobster_tidy, aes(sample= size_mm)) +
  geom_qq() +
  facet_wrap(~site, scales = "free")

```
