---
title: "ESM 206 Lab 5"
author: "David Segan"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridate} package to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}

library(tidyverse)
library(here)
library(janitor)

```


```{r}

# Import lobster data

lobster_abundance <- read_csv(here("data","lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()

```

Use the tidyr::uncount() function to expand a frequency table. Expands observation for as many counts there are listed. Gets rid of count column.

```{r}

lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

Yayayay! Now each lobster has its own row. This is tidy format.

Now let's look at this a bunch of ways.

```{r}

ggplot(data = lobster_tidy,
       aes(x = site,
           y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)

ggplot(data = lobster_tidy,
       aes(x = site,
           y = size_mm)) +
  geom_violin(aes(color = site),
              width = 0.6,
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill=site)) +
  facet_wrap(~site, scales = "free") # scales = free means individual scales can differ based off maxes and mins

ggplot(data = lobster_tidy, aes(sample= size_mm)) +
  geom_qq() +
  facet_wrap(~site, scales = "free")

```


Use 'lubridate' to parse dates and times:

I'm going to add a new column with mutate() that contains my date as an actual date.

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(date_new = lubridate::mdy(date))

```

Parse 'Date' column to get the different pieces (month, year) separated

```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(obs_month = lubridate::month(date_new, label = TRUE), # label = true to have name of month instead of the month number
         obs_year = lubridate::year(date_new))
  
```

Count lobsters by different groupings...

Let's say I want to count the number of lobsters by year and month

```{r}

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_yr_site <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_site <- lobster_parse_date %>% 
  count(site)

# group_by and summarize and n()

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )

```

Confidence intervals and t-tests

Use the 't.test()' function to find confidence intervals and perform t-tests

```{r}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)

# Alternative: t.test(ivee_lobsters$size_mm) # this avoids pull

```

Two-sample t-test:
Is there a significant difference in mean lobster lengths between Naples Reef and Mohawk Reef?

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site =="NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)


mohk_napl_ttest <- t.test(napl_sample, mohk_sample)
mohk_napl_ttest
```

p-value: if samples were drawn from a population with the same mean, this is the probability that the sample means would be so different


Mean lobster size differed significantly between Mohawk and Naples reefs t(`r mohk_napl_ttest$parameter`) = 

Maybe an easier way....

```{r}

lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)

```

Make a geom_tile heatmap

```{r}

ggplot(data = lobster_yr_site, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))

```

